---
title: "R Notebook for predicting LPSA"
output: html_notebook
---
# 1  Exploratory data analysis andKnearest neigh-bor regression

```{r}
prostate=read.table('prostate.data',header = TRUE)
```
## We make sure to visualize the summary of the data

```{r}
summary(prostate)
```
lcavol lweight ibph icp ipsa are quantitive continue (float)
age gleason ipgg are quantitative discrete (integer)
svi and train are booleans

## display using scatter plots and boxplots

```{r}
plot(prostate, label="plot de toutes les datas")
```
```{r}
plot(prostate$lcavol,prostate$lpsa)
```
Plots are good for visualize continus var
```{r}
plot(subset(prostate,select = c(lpsa,lweight,age,lcavol,lbph,lcp)))

```
```{r}
plot(prostate[-10])
```

When vizualize quantitative vs qualitative a boxplot is adapted
```{r}
boxplot(lpsa~svi,data=prostate,xlab="svi",ylab="lpsa")
boxplot(lpsa~gleason,data=prostate,xlab="svi",ylab="gleason")
```
##  predict lpsa 
### we will use lpsafrom input variableslcavol,lweight,ageandlbph and knn algorithm


```{r}
library('FNN')
```

select the interesting rows
```{r}
data= prostate[,c('lcavol','lweight','age','lbph','lpsa','train')]
```

normalize
```{r}
x.train=scale(data[data$train==T,1:4])
y.train=data[data$train==T,5]
x.tst=scale(data[data$train==F,1:4])
y.tst=data[data$train==F,5]
```

use of knn
```{r}
reg<-knn.reg(train=x.train, test = x.tst, y=y.app, k = 4)
```

model error:
```{r}
mean((y.tst-reg$pred)**2)
```
```{r}
plot(y.tst,reg$pred,xlab='y',ylab='prediction')
abline(0,1)#c'est la ligne si les prédictions etaient parfaites
```
## with the graphic we can choose the k so that the error is minimal

```{r}
MSE<-rep(0,15)
for(k in 1:15){
  reg<-knn.reg(train=x.train, test = x.tst, 
               y=y.train, k = k)
  MSE[k]<-mean((y.tst-reg$pred)^2)
}
plot(1:15,MSE,xlab='k',ylab='MSE')  
```
```{r}
which.min(MSE)#find the min index in the list
```

# 2 Normal data generation and knearest neighbor regression

## we generate test and train sets
```{r}
library(mvtnorm)
```
set parameters of the conditional densities P(Y|X) X~N(mu,sigma)
```{r}
mu1=c(0,0)
mu2=c(0,2)
mu3=c(2,0)
```

```{r}
sigma1=matrix(c(1,0.5,0.5,2),2,2)
sigma2=matrix(c(1,0.5,0.5,2),2,2)
sigma3=matrix(c(1,0.5,0.5,2),2,2)
```
funtion that generate a dataset
```{r}
gen.data<- function(N,mu1,mu2,mu3,Sigma1,Sigma2,Sigma3,p1,p2){
  y<-sample(3,N,prob=c(p1,p2,1-p1-p2),replace=TRUE)
  X<-matrix(0,N,2)
  N1<-length(which(y==1)) # number of objects from class 1
  N2<-length(which(y==2))
  N3<-length(which(y==3))
  X[y==1,]<-rmvnorm(N1,mu1,Sigma1)
  X[y==2,]<-rmvnorm(N2,mu2,Sigma2)
  X[y==3,]<-rmvnorm(N3,mu3,Sigma3)
  return(list(X=X,y=y))
}
```
build one training and one test set

```{r}
train<-gen.data(N=100,mu1,mu2,mu3,sigma1,sigma2,sigma3,p1=0.3,p2=0.2)
test<-gen.data(N=1000,mu1,mu2,mu3,sigma1,sigma2,sigma3,p1=0.3,p2=0.2)
plot(train$X[,1],train$X[,2],col=train$y,pch=train$y)
```
## let's classify with knn k=5
```{r}
ypred<-knn(train$X,test$X,factor(train$y),k=5)
```
let's print adjacent matrix
```{r}
table(test$y,ypred)
```
mean error
```{r}
err<-mean(test$y != ypred)
print(err)
```
Initializing matrix that are recording erreur for each [random generation set, k ] with 0s
```{r}
M<-10 #number of random generation of sets
Kmax<-20 

ERR100<-matrix(0,M,Kmax)
ERR500<-ERR100
```
```{r}
for(m in 1:M){
  print(m)
  train100<-gen.data(N=100,mu1,mu2,mu3,sigma1,sigma2,sigma3,p1=0.3,p2=0.2)
  train500<-gen.data(N=500,mu1,mu2,mu3,sigma1,sigma2,sigma3,p1=0.3,p2=0.2)
  for(k in 1:Kmax){
    ypred<-knn(train100$X,test$X,factor(train100$y),k=k)
    ERR100[m,k]<-mean(test$y != ypred)
    ypred<-knn(train500$X,test$X,factor(train500$y),k=k)
    ERR500[m,k]<-mean(test$y != ypred)
  }
}
err100<-colMeans(ERR100)
err500<-colMeans(ERR500)
plot(1:Kmax,err100,type="b",ylim=range(err100,err500))
lines(1:Kmax,err500,col="red")
```

We see that the error tends faster to zero with a bigger trainning set